How to compute the fusion product of MV cycles in type A

### 2021-03-03 draft.tex
Pseudocode and updated examples.

% \setemojifont{Apple Color Emoji}
% Working title: 
\title{How to compute the fusion product of MV cycles in type A}
\author{Roger Bai, Anne Dranowski, Joel Kamnitzer} 
\date{October 2020}

@@

\jcom{Maybe need that both $K$ and $R$ are integral domains and have the same field of fractions?} 
\acom{Roger has proved that what is written is ok.}
% is this true?

% The trivial lattice is 
$ L_0(R) = R^m \subset K^m $ is called the standard lattice. 
We will abuse notation and dentoe $L(R)$ simply $L_0$ in the hope that the ring $R$ will be understood from the context. 
The group $\GL_m(K) $ acts transitively on the set of $R$-lattices in $K^m$, thus giving a bijection between 
% the set of such lattices 
this set and $\GL_m(K)/\GL_m(R)$. 
% \acom{Maybe abuse notation and warn the reader.}
\acom{is it clear that $\GL(R^m) \le \GL(K^m)$ is normal?}

@@

\begin{proof}
First we recall that for $ L \in \Gr^+$, $ L \in \Gr^{\lambda} = G(\CC\xt)t^\lambda $ if and only if $ t |_{L_0/L} $ has Jordan type $ \lambda$.

    Now assume that $ L $ lies in the image of the map from $ \Gr^{\lambda_1, \lambda_2}_s$.  We know that $ L(0) \in \Gr^{\lambda_1}, L(s) \in \Gr^{\lambda_2} $.  This means that $t $ acting on $ \CC\xt^m / L(0)$ has Jordan type $ \lambda_1$ and $ t$ acting on $ \CC\xt[t-s]^m / L(s) $ has Jordan type $ \lambda_2$.  Now, the map $ L_0 / L \rightarrow \CC\xt^m / L(0)$ induces an isomorphism between the $0$-generalized eigenspace of $ t$ and $ \CC\xt^m / L(0)$.    This shows that (i) implies (ii) and the logic can be reversed to see that (ii) implies (i).

@@

\begin{lemma} \label{le:Wmu}
    Let $ L \in \Grth^+$.  The following are equivalent:
    \begin{enumerate}
        \item $ L \in \cW_\mu$.
                \item $ L = \Sp_\cO(v_1, \dots, v_m)$ for some $ v_i $ of the form $ v_i = t^{\mu_i} e_i + \sum_{j=1} p_{ij} e_j $ where $ p_{ij} \in \CC[t] $ has degree less than $ \min(\mu_i, \mu_j)$.
        \item $ \beta_\mu := \{ [t^k e_i] : 0 \le k < \mu_i, 1 \le i \le m\}$ forms a basis for $ L_0/L$.

    \end{enumerate}
\end{lemma}

\begin{proof}
    Let $ L \in \cW_\mu$.  Then $ L = \Sp_\cO(v_1, \dots, v_m) $ for some $ v_i $ with $ v_i = t^{\mu_i} e_i + \sum_{j=1} q_{ij}t^{\mu_i} e_j $ and $ q_{ij} \in t^{-1} \CC\xt[t^{-1}]$.  Since $ L \in \Grth $, we see that $ v_i \in L_0$ which mean that $ p_{ij} := q_{ij}t^{\mu_i} \in \CC[t]$.  If $ \mu_j < \mu_i$, then we can alter this basis to $ v'_i = v_i - r v_j$ for some polynomial $r \in \CC[t]$ and thus ensure that $ p_{ij} $  has degree less than $ \min(\mu_i, \mu_j)$.  Thus (i) implies (ii).

    Suppose that $ L = \Sp_\cO(v_1, \dots, v_m)$ as in (ii).  Then $[t^{\mu_i} e_i] \in \Sp_\CC(\beta_\mu) $
in $ L_0/L$.  For this, we can deduce that $ \beta_\mu$ forms a basis for $ L_0/L$.  Hence (ii) implies (iii).  
    Finally, if we know that $ \beta_\mu $ forms a basis, then $ v_i := t^{\mu_i} e_i - \sum_{j=1} p_{ij} e_j \in L $ for some $ p_{ij} \in \CC[t]$ of degree less than $ \mu_j $.  It is easy to see that $ L = \Sp_\cO(v_1, \dots, v_m) $ and so $ L \in \cW_\mu$.  
    \end{proof}

@@

\begin{proof}
    Let $ L $ lie in this image.  Then $ L = g t^{\mu_1} (t-s)^{\mu_2}$ from some $ g \in N(\CC[t, t^{-1}, (t-s)^{-1}]) $.  Let $ h =(t-s)^{\mu_2} t^{-\mu_2}  $.  Note that $ h \in T_1\CC(\xt[t^{-1}])$. Then,
    $$ L = h (h^{-1} g h) t^{\mu}$$
    Note that $ h^{-1} g h \in N(\CC\xT[t^{-1}])$ and from KWWY \jcom{Check this}, we know that $ (h^{-1} g h) t^{\mu} \in \cW_\mu$ and so the result follows.
    \end{proof}

\section{Climax}

\begin{theorem} 
The map $ \TT_\mu \rightarrow \Grth^+ \cap \cW_\mu $ given by $ A \mapsto g(A) $ is an isomorphism with inverse given by
$$ L \mapsto [t|_{L_0/L} ]_{\beta_\mu}$$
\end{theorem}

\jcom{Do we have a reference?}
\acom{Your paper with Sabin??}
\jcom{I took at a look at Mirkovic-Vybornov (both versions) and at my paper with Sabin.  In Mirkovic-Vybornov, there is nothing about the image of the map.  In the paper with Sabin, we describe the image as $ L \subset L_0$, $[e_1], \dots $ forms a basis and $ z^{\mu_i} e_i \in W_{\mu_i} + L$.  I don't know why I put that third condition.  It seems automatic to me, since from the ``basis'' condition we know that $W_{\mu_i} + L = L_0$.  Anyway, I think that I will try to write out the proof again for this paper.}
\acom{Rather $(\sum W_{\mu_i} )+ L = L_0$? But yes we agree that the basis assumption is enough. Can try to prove by induction?}

@@

\begin{theorem} \label{th:OGrl}
    There is an isomorphism
    $$\overline{\OO}^{\lambda_1, \lambda_2}_\AA \cap T_\mu \cong \Gr^{\lambda_1, \lambda_2}_\AA \cap \cW_\mu $$
    given by $ (A,s) \mapsto (g(A), s)$.
\end{theorem}

>>>>

% \setemojifont{Apple Color Emoji}
% Working title: 
\title{How to compute the fusion product of MV cycles in type A}
\author{Roger Bai, Anne Dranowski, Joel Kamnitzer} 
% \date{October 2020}
\date{\today}

@@

\jcom{Maybe need that both $K$ and $R$ are integral domains and have the same field of fractions?} 
\acom{Roger has proved that what is written is ok.}
\rcom{Here's my argument: If $L \subset K^m$ is a free rank $m$ $R$-module and $L \otimes K = K^m$, then it's clear we that $L$ is generated by $m$ generators $v_1,\dots,v_m$ over $R$. Since $L\otimes K = K^m$, these also generate $K^m$ over $K$. Then we have a surjective map $K^m \twoheadrightarrow K^m$ given by sending a basis of $K^m$ to the generators $v_1,\dots,v_m$. Since any surjective endomorphism of finite $K$-modules is injective, the $v_1,\dots,v_m$ form a basis. I think the only thing that is needed is that the rings are commutative.}
% is this true?

% The trivial lattice is 
$ L_0(R) = R^m \subset K^m $ is called the standard lattice. 
We will abuse notation and dentoe $L(R)$ simply $L_0$ in the hope that the ring $R$ will be understood from the context. 
The group $\GL_m(K) $ acts transitively on the set of $R$-lattices in $K^m$, thus giving a bijection between 
% the set of such lattices 
this set and $\GL_m(K)/\GL_m(R)$. 
% \acom{Maybe abuse notation and warn the reader.}
% \acom{is it clear that $\GL(R^m) \le \GL(K^m)$ is normal?} \jcom{It isn't normal and the quotient is not a group.}

@@

\begin{proof}
First we recall that for $ L \in \Gr^+$, $ L \in \Gr^{\lambda} = G(\cO)t^\lambda $ if and only if $ t |_{L_0/L} $ has Jordan type $ \lambda$.

    Now assume that $ (L, s) \in \Gr^{\lambda_1, \lambda_2}_s$.  We know that $ L(0) \in \Gr^{\lambda_1}, L(s) \in \Gr^{\lambda_2} $.  This means that $t $ acting on $ \CC\xt^m / L(0)$ has Jordan type $ \lambda_1$ and $ t$ acting on $ \CC\xt[t-s]^m / L(s) $ has Jordan type $ \lambda_2$.  Now, the map $ L_0 / L \rightarrow \CC\xt^m / L(0)$ induces an isomorphism between the $0$-generalized eigenspace of $ t$ and $ \cO^m / L(0)$.   The same thing holds for the $s$-generalized eigenspace of $t $ and $ \cO_s^m/L(s)$. This shows that (i) implies (ii) and the logic can be reversed to see that (ii) implies (i).

@@

\begin{lemma} \label{le:Wmu}
    Let $ L \in \Grth^+$.  The following are equivalent:
    \begin{enumerate}
        \item $ L \in \cW_\mu$.
                \item $ L = \Sp_\cO(v_1, \dots, v_m)$ for some $ v_i $ of the form $ v_i = t^{\mu_i} e_i + \sum_{j=1} p_{ij} e_j $ where $ p_{ij} \in \CC[t] $ has degree less than $ \min(\mu_i, \mu_j)$.
        \item $ \beta_\mu := \{ [t^k e_i] : 0 \le k < \mu_i, 1 \le i \le m\}$ forms a basis for $ L_0/L$ and for all $ i $, 
        $$ t^{\mu_i} e_i \in \Sp_\CC(\{t^k e_j : 0 \le k < \min(\mu_i, \mu_j), 1 \le j \le m \}) + L $$.
    \end{enumerate}
\end{lemma}

\jcom{I realized the meaning of the ``extra'' condition from the paper with Sabin and I added it into point 3.}

\begin{proof}
    Let $ L \in \cW_\mu$.  Then $ L = \Sp_\cO(v_1, \dots, v_m) $ for some $ v_i $ with $ v_i = t^{\mu_i} e_i + \sum_{j=1} q_{ij}t^{\mu_i} e_j $ and $ q_{ij} \in t^{-1} \CC\xt[t^{-1}]$.  Since $ L \in \Grth^+ $, we see that $ v_i \in L_0$ which means that $ p_{ij} := q_{ij}t^{\mu_i} $ lies in $ \CC[t]$.  By construction, the polynomial $ p_{ij}$ has degree less than $ \mu_i$.

    Fix $ i$ and suppose that for some $ j$, $ \mu_j < \mu_i$.  In this case, we can alter our basis to $ v'_i = v_i - r v_j$ for some polynomial $r \in \CC[t]$.  This gives us new polynomials $ p'_{ij} = p_{ij} - r (t^{\mu_j} + p_{jj}) $.  In this way, we can ensure that $ p_{ij} $ has degree less than $ \min(\mu_i, \mu_j)$.  Thus (i) implies (ii).

    Suppose that $ L = \Sp_\cO(v_1, \dots, v_m)$ as in (ii).  Then $[t^{\mu_i} e_i] \in \Sp_\CC(\beta_\mu) $
in $ L_0/L$ and we see that 
$$t^{\mu_i} e_i \in \Sp_\CC(\{t^k e_j : 0 \le k < \mu_i \}) + L $$.
From this, we can deduce that $ \beta_\mu$ forms a basis for $ L_0/L$.  Hence (ii) implies (iii).  
    Finally, given (iii), then we can see $ v_i := t^{\mu_i} e_i - \sum_{j=1} p_{ij} e_j \in L $ for some $ p_{ij} \in \CC[t]$ of degree less than $ \min(\mu_i,\mu_j) $.  It is easy to see that $ L = \Sp_\cO(v_1, \dots, v_m) $ and so $ L \in \cW_\mu$.  
    \end{proof}

@@

\begin{proof}
    Let $ L $ lie in this image.  Then $ L = g t^{\mu_1} (t-s)^{\mu_2}$ from some $ g \in N(\CC[t, t^{-1}, (t-s)^{-1}]) $.  Let $ h =(t-s)^{\mu_2} t^{-\mu_2}  $.  Note that $ h \in T_1(\CC(\xt[t^{-1}]))$. Then,
    $$ L = h (h^{-1} g h) t^{\mu}$$
    Note that $ h^{-1} g h \in N(\CC\xT[t^{-1}])$ and from KWWY \jcom{Check this}, we know that $ (h^{-1} g h) t^{\mu} \in \cW_\mu$ and so the result follows.
    \end{proof}

\section{Climax}
Given $ A \in \TT_\mu$, recall the definition of $ g(A)$ given in section ??.  As $ g(A) \in M_n(\CC[t]) \cap G_1(\CC\xt[t^{-1}])$, we will regard $ g(A)$ as giving an element of $ \Grth^+ \cap \cW_\mu$.

The following result in \cite[Theorem 3.2]{cautis2018categorical}.

\begin{theorem} \label{th:TmuWmu}
The map $ \TT_\mu \rightarrow \Grth^+ \cap \cW_\mu $ given by $ A \mapsto g(A) $ is an isomorphism with inverse given by
$$ L \mapsto [t|_{L_0/L} ]_{\beta_\mu}$$
\end{theorem}

@@

\begin{theorem} \label{th:OGrl}
    There is an isomorphism
    $$\overline{\OO}^{\lambda_1, \lambda_2}_\AA \cap \TT_\mu \cong \Gr^{\lambda_1, \lambda_2}_\AA \cap \cW_\mu $$
    given by $ (A,s) \mapsto (g(A), s)$.
\end{theorem}

\begin{proof}
Since we already have the isomorphism from Theorem \ref{th:TmuWmu}, it suffices to show that for any $ A \in \TT_\mu$, $$ (A,s) \in \overline{\OO}^{\lambda_1, \lambda_2}_\AA \cap T_\mu \text{ if and only if } (g(A), s) \in \Gr^{\lambda_1, \lambda_2}_\AA \cap \cW_\mu $$
This follows immediately from Lemma \ref{le:Grl1l2}.
\end{proof}
